An Innovative Storage Stack Addressing Extreme Scale Platforms and Big Data Applications

### 2014-05-12 paper.tex
removed the large commented out sections
spell check completed. Initial submission version.

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase 1 of the project complete, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper examimes the overall design.
Each layer of the proposed stack is discussed in some detail along with
cross-cutting topics that impact multiple layers, such as transactions and
metadata management. We also describe the proposed features to better support
some classes of big data applications.

This paper not only provides a timely summary of important aspects of the
design specifications but also captures the underlying reasoning that is not
available elsewhere. With this information, we encourate the broader community
to understand the design, intent, and future directions to foster discussion
guiding phase 2 and the ultimate production storage stack based on this work.
An initial performance evaluation of the early prototype implementation is also
provided to validate the presented design.

@@

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack (in green) and offers features like Burst
Buffers to insulate the persistent storage array from bursty IO workloads.
Ideally, the IOD layer's functionality can be optional based on available
hardware and compute power provided on the IO Nodes (IONs). Much of the
functionality offered at this layer would shift either up or down the stack as
discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
layer serves as the persistent storage interface and translation layer between
the user-visible object model and the requirements of the underlying storage
infrastructure. It is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albiet with the potential performance penalty of manipulating
the shared, persistent storage array. At the bottom is the Versioning Object
Storage Device (VOSD) (in purple).  It serves as the interface for storing
objects of all types efficiently for each storage device in the parallel
storage array. Think of this layer as the physical disk interface layer. In
terms of Lustre, this would replace the API on individual storage devices with
an interface friendlier to the containers of objects and transactions/epochs
concepts used in the higher layers.

@@

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the contect of the HDF5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
layer functions. As with the IOD layer, the cross-cutting features will be
mentioned, but disucssed more fully in the cross-cutting section. The VOSD
layer is disucssed in Section~\ref{sec:vosd}. In particular, the mapping
between the DAOS and VOSD layers are explored as it pertains to the physical
storage. Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}. Since these and other
features are spread across multiple layers, it makes more sense to discuss them
independently once an understanding of the overall structure has been
presented.  A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}. This shows that the prototype system based on the
proposed design can function. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

Lustre~\cite{braam:lustre-arch} has become the defacto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.

@@

%\item
%  Virtual Object Layer (VOL) offers the ability to plug in new implementations
%  for HDF5 semantics. In this case, the HDF5 calls are implemented using the
%  IOD or DAOS API rather than MPI-IO or POSIX.

@@

%\begin{figure}[htbp]
%\vspace{-0.10in}
%\centering
%\includegraphics[width=\columnwidth]{images/hdf5.png}
%\vspace{-0.20in}
%\caption{Sample file illustrating the hierarchical structure of HDF5 files}
%\label{fig:hdf5-sample}
%\vspace{-0.25in}
%\end{figure}

@@

%For example, as shown in Figure \ref{fig:hdf5-sample}, the file Sample.h5
%contains the root group which itself contains a group G1 and two datasets, D1
%and D2. Group G1 contains a dataset D3. Attribute A1 is linked to dataset D1.
%The objects and the relationships between them can be represented as a B-tree,
%which is used internally by HDF5 to index its objects.

@@

%\subsection{IOD VOL plugin}
%\label{iod-vol-plugin}

@@

For the basic HDF5 calls, this will work the same as how the Nessie
staging~\cite{lofstead:2011:nessie-staging} shifted the collective IO data
rearrangement calls to a reduced number of processes. The prototype
implementation will only support accessing functionality already deployed to
the IO nodes through an RPC mechanism. This initial implementation will use
Mercury~\cite{Soumagne:2013:mercury} to access the remote functionality. For
dynamically defined functions, a different system will be required leveraging
something like C-on-demand~\cite{abbasi:2011:c-on-demand} or some other dynaimc
deployment and compilation or an interpreter system.

@@

Strictly speaking, the IO Dispatcher layer and included functionality, such as
burst buffers, is optional. All of the functionality, such as function and
analysis shipping, transaction management, and managing asynchronous data
movement can be handled by other portions of the stack. For an extreme scale
platform, the IOD layer will be an essential pressure relief valve for the
underlying persistent storage layer. By making it optional, the proposed stack
can be deployed more easily on smaller clusters or for those with more
constrainted budgets. For simplicity, the rest of this section will describe a
full stack including all of the proposed IOD components.

@@

%\subsection{IO Dispatcher API}
%\label{sec:fast-forwards-io-dispatcher-api}
%
%The Fast Forward IO and Storage project is aimed at integrating many of
%the services existing currently in middleware libraries in a
%next-generation storage stack, exposed through a non-POSIX,
%transactional, asynchronous, object-based API.
%
%From a high-level point of view, the Fast Forward's IO Dispatcher (IOD)
%interface allows the user to:
%
%\begin{itemize}
%\itemsep1pt\parskip0pt\parsep0pt
%\item
%  transactions
%\item
%  asynchrony
%\item
%  objects
%\item
%  layout
%\item
%  placement
%\item
%  formats
%\end{itemize}
%
%Type of objects:

@@

Attributes are stored in a key-value store object, but use the
multi-dimnensional array and blob objects to store the values for the
attributes.

@@

For both IO performance and to aid in analysis and other data processing, the
multi-dimensional array object can be split across multiple IO nodes. Each
piece of this array is called a {\em shard}.  The idea of sharding is to store
a logically complete portion of a data set on a single storage target. This is
similiar in concept to the HDF5 hyperslab.  The FFSIO stack supports sharding
the data in the default or some other structured way as well as ``re-sharding''
based on application needs. For example, reordering the data so that a
different dimension is the ``fast'' dimension may greatly improve the
performance of a subsequent data analytics task. A common scenario where this
is common is a Fortran code (column-major) writes data for a C code (row-major)
to analyze.  The IOD API supports the following sharding strategies:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{contiguous}. fixed chunking, distributed in a round-robin
  fashion accross the IO nodes.
\item
  \textbf{chunked}. same as above but with irregular (sparse) chunking.
\item
  \textbf{user-defined}. either contiguous or chunked, but user specifies
  where to place each individual shard.
\end{itemize}

@@

All of these distinct alternatives result in having many different ways for
executing the same analysis task.
%In the future, we plan to extend our
%design to incorporate an optimization phase that will prepare a task by
%taking into account the distinct possible ways of executing and select
%the most promising one.
In the subsequent discussions, we consider only
data-movement optimization, i.e., sending the analysis code as close as
possible to the data. In practice, this means we focus on identifying
sharding of datasets and execute code accordingly over the appropriate
shards.

%At the DAOS layer, the sharding can be changed based on the type of storage
%devices beneath it.

@@

\begin{figure}[htbp]
\centering
\vspace{-0.10in}
\includegraphics[width=\columnwidth]{images/exa-arch.png}
\vspace{-0.20in}
\caption{Exascale Architecture}
\label{fig:exa-arch}
\vspace{-0.10in}
\end{figure}

A common configuration for this type of deployment is shown in
Figure~\ref{fig:exa-arch}. The designated IO nodes (IONs) are connected to the
compute nodes (CNs) through the same fast fabric (e.g., InfiniBand) while the
connection to the external storage cluster is through a secondary, slower
channel (e.g., 10Gb Ethernet). By providing additional storage on the IO nodes,
such as SSDs, these nodes are capable of better regulating the IO pressure on
the underlying storage array better than simple forwaring gateways. For this
project, using something like SSDs on the IO nodes is termed a {\em Burst
Buffer} and is disucssed below.

%In current proposals, the stack is managed by middleware that sits between the
%application and the parallel file system. Since most of the applications assume
%a POSIX interface to storage, existing middleware either transparently handle
%the IO operations, appearing as regular POSIX calls to applications; or modify
%the IO API as little as possible in order to minimize the impact on production
%codes.
%
%As we move towards the exascale goal, this way of interfacing with storage
%becomes more of an obstacle. Many features provided by the stack are hidden by
%the middleware/POSIX layers. If these were visible to the applications, and if
%the applications were able to control how the underlying layers behave, they
%would benefit greatly since the domain knowledge that is available at the
%application level could be used to execute storage operations efficiently. One
%more implication of the 25-year-old POSIX API is in cumbersome middleware,
%since metadata has to be part of the same uni-dimensional byte array being
%stored.
%
%The Fast Forward Storage IO project is aimed at merging the features of
%existing middleware into a next-generation parallel file system stack. This new
%storage API termed the IO dispatcher (IOD) is object-based, transactional,
%asynchronous and IO-staging-aware (Figure~\ref{fig:exa-stack} (right)).

%\begin{figure}[htbp]
%\centering
%\includegraphics[width=\columnwidth]{images/exa-stack.png}
%\caption{Top-to-bottom view of the component stack}
%\label{fig:exa-stack}
%\end{figure}

%The prototype stack, from a top-down point of view, has the following
%components:
%
%\begin{enumerate}
%\def\labelenumi{\arabic{enumi}.}
%\itemsep1pt\parskip0pt\parsep0pt
%\item
%  Arbitrarily Connected Graphs (ACG).
%\item
%  HDF5 extensions.
%\item
%  IO Dispatcher (IOD).
%\item
%  Distributed Asynchronous Object Storage (DAOS).
%\item
%  Versioning Object Storage Device (VOSD).
%\end{enumerate}
%
%Applications running on the compute nodes will be written in Python scripts or
%in C++ using the HDF5 APIs. These applications will execute analysis on
%graph-based data (ACG). The point is to demonstrate both the HPC and BigData
%use cases of the Exascale architecture. The data structures will be stored in
%HDF5, for which new API extensions will be implemented. These new API calls
%will expose the transactional, asynchronous and function-forwarding semantics
%of the underlying stack.

%As mentioned before, the stack provides \emph{non-POSIX, object-based,
%transactional and asynchronous active-storage}, meaning that POSIX is
%supplanted by new object interfaces that reach up to the HDF5 layer
%(Application IO in the stack diagram); transactional semantics are present,
%from HDF down to the VOSD layer (Storage layer); clients don't need to wait for
%any blocking operation; and analysis can be shipped and executed on IO or DAOS
%nodes.

@@

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any dedicated hardware support for this functionality.
The desired outcome of these initial studies is to motivate how such
functionality might be incorporated and the potential benefits.  Later, these
concepts were proposed to be incorporated into the existing IO stack
architecture~\cite{nowoczynski:2008:zest,bent:2012:challenges,bent:2012:burst-buffer}.
%The current Fast Forward IOD design recommends incorporating SSDs, but
%specifically lists these devices as optional. Unfortunately, not incorporating
%burst buffers and the use of SSDs in the IOD layer may be problematic.  First,
%the IOD design currently is written assuming burst buffers. This means that the
%bulk of IO operations will only hit the IOD layer and proposed functionality,
%such as the function shipping, do not discuss the impact on the DAOS layer or
%the operators themselves should a burst buffer not be available.  Consider the
%important functionality of data rearrangement and doing things like changing
%the fast array dimension on shared, spinning media.  The design of DAOS assumes
%that it will only be involved when persisting a completed transaction and only
%for a fraction of the total transactions created.  Transaction frequency at the
%IOD layer can be higher since IOD does not ``flatten'' transactions. Flattening
%is the process of transferring the set of copy-on-write changes from
%transaction/epoch $n$ to transaction $n+m$ into the DAOS layer. The DAOS layer
%mirrors the copy-on-write functionality, but stores the differences between
%epochs rather than individual transactions. The primary difference is that at
%the DAOS layer, data is stored in large, logically contiguous chunks to manage
%the metadata load and hopefully improve read performance.

%One of the bigger concerns is the observation that the original data staging
%proposals all used compute nodes while the newer proposals seek not only to
%make them a fixed portion of the IO stack, but also shared across all machine
%users. The PreDatA~\cite{zheng:2010:predata} paper in particular examines the
%potential costs and advantages of where to place operators similar to the IOD
%proposed function shipping. There are two key takeaways from PreDatA. First,
%placement matters.  Depending on the communication intensity vs. computation
%intensity, where along the IO path to place the operation can matter
%significantly. Second, and more importantly, the amount of time spent
%processing for the operators was stretched to the point where it consumed
%nearly all of the time between IO operations. The given ratios of compute
%processes to staging process examined is representative for future extreme
%scale platforms. If anything, the ratios offer more staging processes than IOD
%processes would be available.
%

@@

%This is unlikely to be useful
%because of the limited compute and communication capacity to spare to perform
%these operations at a bottleneck in the IO path.  The use of a separate data
%staging area intentionally separate from the IO path allows using operators on
%limited resources leaving the IO path clear for strictly data movement. A
%nuance of this design is discussed in the Design Philosophy below.

@@

%By concentrating the Burst Buffers and function shipping into the storage
%stack, three problems arise.  First, the amount of network bandwidth, IO
%bandwidth, and compute power consumed for example operations from a single
%application is likely to completely monopolize the IOD processes. Second, if
%space and time partitioning is used instead, the functionality risks being too
%small to be useful. Third, the long-term hardware performance advantage for
%SSDs is questionable. Recent studies have shown that the erase-before-write
%and interference between reading and writing with flash-based SSDs can cause
%severe performance problems~\cite{skourtis:2013:ssd-performance}. The inclusion
%of an optional use SSD layer in the new Trinity machine at Los Alamos will
%offer a test bed to determine how likely these observed problems would affect a
%production extreme scale platform.
%
%Given these features, the optionality and even incorporation of burst buffers
%in the current design should be carefully considered. Much of the advanced, key
%functionality proposed as they are currently designed ultimately relies on the
%existence of burst buffers to work. Further thought about how to have an IOD
%layer both with and without a burst buffer is required before they can be
%considered optional. As the design stands today, they are a required part of
%the IOD layer for proper functioning. Unfortunately, it is not clear that they
%can address the performance concerns they are intended to cover.

@@

%Another concern that is acknowledged, but no thought has been applied to, is
%the requirement that a single IOD process of the set assigned to an application
%is the master for any operation.  Should the number of concurrent applications
%exceed the available nodes, sharing an IOD process will be required. The
%requirements both in terms of scheduling and resource management were cut from
%the project due to funding limitations. Since partitioning of the IOD processes
%for exclusive use by particular applications is the assumed operating mode,
%should insufficient IOD resources be available, either a job could be delayed
%or IOD resources could be reallocated from a different process could be
%redeployed for use by the new job. Handling resilience concerns for the IOD
%processes must also be address.  These sorts of considerations still need to be
%made for a full production system.

@@

This is the layer where the container/object model is translated into the
physical storage requirements dictated by the physical storage underneath (the
VOSD layer). The two key design elements of this layer are the handling of
epochs and the mapping of conatiners and objects to the underlying storage.

@@

The current implementation has the DAOS layer map the container/object data
model onto a directory/file data model used for most existing file systems.
Should a fully object-based file system be deployed at the VOSD layer, this
mapping would be unnecessary. The current projections suggest that a standard
POSIX-like file system will likey be used at the lowest level on each storage
device requiring the mapping at some level. To perform this mapping, DAOS
considers the following.

@@

%Epochs are represented through the shard mechanism. Each container shard is
%versioned according to the transactions in the IOD layer that are persisted to
%the DAOS layer as an epoch. The container shard is a piece of the container
%with the contained objects concept. Since the container is largely logical,
%this generally maps to a shard of an object within a container.

@@

At a more detailed level, the design for VOSD is an increment beyond the
current Lustre Object Storage Device design to incorporate the idea of shards
and the versioning aspects of transactions/epochs. For every DAOS shard, the
VOSD has information for storing and accessing the currently commited version,
the Highest Committed Epoch, as well as a staging dataset represeting the
next version of the object being stored. Both of these are combined in a shard
root.

For data integrity, an intent log is maiained as part of the underlying file
system enabling fault recovery.

@@

In the subsections that follow, we examine transactions and epochs, metadata
management, function and analysis shipping, and the aribtrary connected graphs
support.

%At a broader level, there are some concerns that were partially clarified
%through conversations with the team.  Consider a shared file system across an
%HPC data center. The current design maintains the metadata in its own
%container. Since copying data from the IOD layer to the DAOS layer requires an
%explicit persist call, how and when synchronizing the metadata across the
%layers and potentially across machines occurs is undefined. Delaying
%synchronization until an explicit persist is called will reduce the update
%frequency, but delays the data visibility on other platforms. Ideally, the
%metadata object would need to be automatically persisted every time a container
%transaction is persisted to the DAOS layer.
%
%The implication of this is that every transaction persist is double operation
%to account for the metadata persist. More importantly, the IOD-layer version of
%the metadata container may contain readable transactions that have not been
%persisted to the DAOS layer. How to handle this inconsistency between the two
%layers still needs to be explored.
%
%A point of confusion rather than a potential design challenge is the change in
%definitions between the IOD layer and the DAOS layer.  For the IOD layer, a
%container is a collection of objects. For the DAOS layer, a container is a
%collection of objects across a set of shards. For the IOD an object may be a
%shard of a global array.  For DAOS, a shard can host a set of DAOS objects.
%Having the same names with locally correct, globally conflicting definitions
%serves to confuse how the system should work.

@@

As mentioned above, the transaction mechanism manifests in two forms. From the
user level down through the IOD layer, they are called transactions and are
used to judge whether or not a set of distributed, asynchronous modifications
across a set of related objects is complete or not.  It is also used to control
access by treating the transaction ID of committed transaction as a version
identifier.  At the DAOS layer and below, they are called epochs and represent
persisted (durable) transactions from the IOD layer. Each of these offers
different functionality, but are connected as is explained below.
%How these differ from the \DDT approach is also explored.  While IOD's and
%\DDTns's transactions are seemingly very different, they use a similar
%high-level design, but very different implementation, to solve the same
%problem.

@@

\subsubsection{Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted
container copies.  To simplify the mapping between an IOD transaction and the
DAOS epochs, when an IOD transaction is persisted to DAOS, the IOD transaction
ID is the used as the epoch ID. The key difference is that at the DAOS layer,
some transaction (epoch) IDs will not be represented with data since not all
IOD transactions are necessarily persisted. Maintaining this ID continuity is
critial for multiplatform use. Since the shared point is the DAOS layer, any
user adding a new version to a file must be able to determine the most recent
transaction ID no matter from where the container was updated last.

\subsection{Metadata Management}
Metadata management has been a perennial challenge for parallel storage
systems.  Eliminating metadata management as a special case and instead
treating it just as data is a central design goal of the Fast Forward project.
This is a hybrid approach to metadata management that is half-way between
providing no inherent metadata support and having a fully integrated, but
separate metadata management system.

@@

%Based on the lessons from the \DDT metadata
%service~\cite{lofstead:2012:txn-metadata} construction and the prior
%experiments with LWFS, having a completely separate metadata service is
%workable. Rather than making it a bottleneck in the IO path, it is another
%service that users must interact with if they need those services.  Users can
%manage everything by maintaining the metadata including the list of objects
%themselves. However, there are drawbacks to this 100\% client-side approach.
%
%With a client-side only approach, there is a serious risk of the metadata
%service and the object store getting out of sync.  While having a metadata-less
%object storage service is desirable, the different semantics from traditional
%file systems requires some considerations. In this case, should these services
%get out of sync, three particular risks are introduced.  First, a client could
%create a dangling entry in the metadata service that does not correspond to any
%objects in the object store. Second, a client could create orphaned objects
%that have no associated metadata entries. Third, updates to the metadata or
%object store service should be an atomic operation, but due to a lack of
%coordination, a window where the system is inconsistent appears.
%
%Ultimately, the consistency semantics required must be determined. If a
%metadata service is required and it must be in sync with the object storage
%service, then additional work must be performed. In traditional file systems,
%the metadata and object storage updates are atomic. With decoupling metadata
%from object storage, should this atomicity still be desired, it requires both
%the ability for the services to participate in a task that is part of a larger
%atomic operation and a higher-level mechanism to manage the atomic operation.
%
%Overall, while additional work is required to maintain a client-side only
%metadata service, it eliminates any potential bottlenecks related to updating
%metadata related to the object storage. The burden of tracking striping and
%other metadata that has traditionally been part of the metadata associated with
%the file system will have to be maintained by the object storage service. The
%lack of a centralized, serialized bottleneck to store that information improves
%concurrency.

%\subsection{Comparison to \DDTns}
%The \DDT project~\cite{lofstead:2012:txn} sought to develop an efficient
%approach for handling ACID-style transactions in an environment with parallel
%clients and multiple servers (doubly distributed). Rather than being aimed
%solely at data movement operations, \DDT seeks to address the general problem
%of managing any operation with multiple clients and servers.  Consider the
%management of the analysis/visualization area, potentially similar to the IOD
%concept. The transaction protocol is used to help manage resizing of the
%resource allocation to the various analysis and visualization components.  For
%the purposes of this discussion, \DDT could also be used to manage changing how
%IOD processes and/or nodes are used without exposing these changes to the
%client processes prematurely.  This has been described and analyzed
%previously~\cite{dayal:2013:io-containers}.
%
%The example metadata and data storage services created as part of the \DDT
%project have no dependencies between transactions that prevent visibility
%should an older version be incomplete. This additional, intentional requirement
%by IOD offers different functionality than \DDTns's example services. In the
%case of \DDTns, the functionality is more minimal, but also avoids some of the
%concerns outlined below.
%
%The second iteration of the protocol~\cite{lofstead:2013:pdsw-txn} fixed
%scalability issues and demonstrated a scalable client-side coordination model
%with excellent performance. The performance measured for a complex transaction
%with \DDT is illustrated in Figure~\ref{fig:performance}. This performance is
%explored in detail in a previous paper~\cite{lofstead:2013:pdsw-txn}.
%
%\begin{figure}[ht]
%%\vspace{-0.15in}
%\centering
%\includegraphics[keepaspectratio=true, width=0.9\columnwidth]{images/performance}
%\vspace{-0.15in}
%\caption{Total Transaction Overhead}
%\label{fig:performance}
%%\vspace{-0.15in}
%\end{figure}
%
%At a high level, both \DDT and the IOD transactions have the same design. In
%both cases, a hierarchical model is employed. In the case of \DDTns, it is a
%purely client-side tree using semi-synchronous messaging. The messaging itself,
%in the current implementation, uses asynchronous MPI messages. The synchronous
%component comes from the timeout mechanism used to detect faults.  It forces a
%level of coordination and synchronization for the protocol. For IOD, it is a
%server-side tree and fully asynchronous relying on the existing Lustre fault
%detection mechanism for failure detection. In both cases, there is a master in
%charge of managing the transaction and a collection of workers that aggregate
%into the master through second-level leaders. Beyond that, there are some
%significant differences. Some of the different choices made by IOD raise some
%possible concern.
%
%The multi-leader model introduces the possibility of forcing a rollback of an
%entire transaction when a partial retry might be sufficient for success. Since
%the transactions are managed at a high level rather than the individual
%tasks, a failure in a limited distributed task can cause the entire transaction
%to fail. For example, consider 10 processes each have 5 tasks, but 3 of those
%10 have an additional shared task to complete. If the task shared by the 3
%processes fails on any of the three, the entire transaction would roll back
%because it is a coarse-grained success/failure. If a concept like
%sub-transactions at a task granularity were used, then it would be possible for
%the one process that failed to report just that failure. Then the transaction
%manager could reassign the resources for these 3 processes and try just that
%operation again. If it now succeeds, then the overall transaction can be marked
%successful only redoing the minimum amount of work required.
%
%\begin{figure}[ht]
%\centering
%\includegraphics[keepaspectratio=true, width=0.5\columnwidth]{images/optimized-protocol}
%\vspace{-0.15in}
%\caption{Optimized Protocol}
%\label{fig:optimized-protocol}
%%\vspace{-0.25in}
%\end{figure}
%
%\DDT has addressed theses issues in a couple of ways. First, the
%sub-coordinators each have a list of processes from which they expect messages.
%Should a message be missed, it is noticed and corrective action can be taken.
%Second, \DDT has the concept of sub-transactions. The messaging requirements

>>>>

\begin{abstract}
The DOE Extreme-Scale Technology Acceleration Fast Forward Storage and IO Stack
project is going to have significant impact on storage systems design within
and beyond the HPC community. With phase 1 of the project complete, it is an
excellent opportunity to explore the complete design and how it will address
the needs of extreme scale platforms.  This paper examines the overall design.
Each layer of the proposed stack is discussed in some detail along with
cross-cutting topics that impact multiple layers, such as transactions and
metadata management. We also describe the proposed features to better support
some classes of big data applications.

This paper not only provides a timely summary of important aspects of the
design specifications but also captures the underlying reasoning that is not
available elsewhere. With this information, we encourage the broader community
to understand the design, intent, and future directions to foster discussion
guiding phase 2 and the ultimate production storage stack based on this work.
An initial performance evaluation of the early prototype implementation is also
provided to validate the presented design.

@@

Below the user API is an IO forwarding layer that redirects IO calls from the
compute nodes to the IO dispatching layer (in black).  This IO forwarding layer
is analogous to the function of the IO nodes in a BlueGene machine or the
passive data staging processes demonstrated
previously~\cite{nisar:2008:staging,Abbasi:2009:datatap}. The next two layers
have considerable functionality. The IO dispatcher (IOD) serves as the primary
storage interface for the IO stack (in green) and offers features like Burst
Buffers to insulate the persistent storage array from bursty IO workloads.
Ideally, the IOD layer's functionality can be optional based on available
hardware and compute power provided on the IO Nodes (IONs). Much of the
functionality offered at this layer would shift either up or down the stack as
discussed in detail below. The Distributed Asynchronous Object Storage (DAOS)
layer serves as the persistent storage interface and translation layer between
the user-visible object model and the requirements of the underlying storage
infrastructure. It is intended to be the traditional file system-like
foundation on which everything else is built with no dependence on any
technologies specified above it (in dark pink and yellow). For example, the IOD
layer with or without burst buffers is not required for DAOS to operate
properly.  Instead, the DAOS layer can handle all of the IO operations from the
user API layer, albeit with the potential performance penalty of manipulating
the shared, persistent storage array. At the bottom is the Versioning Object
Storage Device (VOSD) (in purple).  It serves as the interface for storing
objects of all types efficiently for each storage device in the parallel
storage array. Think of this layer as the physical disk interface layer. In
terms of Lustre, this would replace the API on individual storage devices with
an interface friendlier to the containers of objects and transactions/epochs
concepts used in the higher layers.

@@

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array. This will be discussed in the context of the HDF5 based
example library used for the functionality demonstration. Section~\ref{sec:iof}
briefly discusses the motivation and proposal for the IO forwarding layer.
Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
functionality it offers. This will detail the pieces of the layer that are
potentially optional and mention the cross-cutting features discussed in a
later, cross-cutting section. Section~\ref{sec:daos} discusses how the DAOS
layer functions. As with the IOD layer, the cross-cutting features will be
mentioned, but discussed more fully in the cross-cutting section. The VOSD
layer is discussed in Section~\ref{sec:vosd}. In particular, the mapping
between the DAOS and VOSD layers are explored as it pertains to the physical
storage. Next is an exploration of cross-cutting features like transactions and
metadata management in Section~\ref{sec:broader}. Since these and other
features are spread across multiple layers, it makes more sense to discuss them
independently once an understanding of the overall structure has been
presented.  A demonstration of the functionality is presented in
Section~\ref{sec:evaluation}. This shows that the prototype system based on the
proposed design can function. The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

@@

Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.

@@

@@

@@

@@

@@

For the basic HDF5 calls, this will work the same as how the Nessie
staging~\cite{lofstead:2011:nessie-staging} shifted the collective IO data
rearrangement calls to a reduced number of processes. The prototype
implementation will only support accessing functionality already deployed to
the IO nodes through an RPC mechanism. This initial implementation will use
Mercury~\cite{Soumagne:2013:mercury} to access the remote functionality. For
dynamically defined functions, a different system will be required leveraging
something like C-on-demand~\cite{abbasi:2011:c-on-demand} or some other dynamic
deployment and compilation or an interpreter system.

@@

Strictly speaking, the IO Dispatcher layer and included functionality, such as
burst buffers, is optional. All of the functionality, such as function and
analysis shipping, transaction management, and managing asynchronous data
movement can be handled by other portions of the stack. For an extreme scale
platform, the IOD layer will be an essential pressure relief valve for the
underlying persistent storage layer. By making it optional, the proposed stack
can be deployed more easily on smaller clusters or for those with more
constrained budgets. For simplicity, the rest of this section will describe a
full stack including all of the proposed IOD components.

@@

@@

Attributes are stored in a key-value store object, but use the
multi-dimensional array and blob objects to store the values for the
attributes.

@@

For both IO performance and to aid in analysis and other data processing, the
multi-dimensional array object can be split across multiple IO nodes. Each
piece of this array is called a {\em shard}.  The idea of sharding is to store
a logically complete portion of a data set on a single storage target. This is
similar in concept to the HDF5 hyperslab.  The FFSIO stack supports sharding
the data in the default or some other structured way as well as ``re-sharding''
based on application needs. For example, reordering the data so that a
different dimension is the ``fast'' dimension may greatly improve the
performance of a subsequent data analytics task. A common scenario where this
is common is a Fortran code (column-major) writes data for a C code (row-major)
to analyze.  The IOD API supports the following sharding strategies:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{contiguous}. fixed chunking, distributed in a round-robin
  fashion across the IO nodes.
\item
  \textbf{chunked}. same as above but with irregular (sparse) chunking.
\item
  \textbf{user-defined}. either contiguous or chunked, but user specifies
  where to place each individual shard.
\end{itemize}

@@

All of these distinct alternatives result in having many different ways for
executing the same analysis task.  In the subsequent discussions, we consider
only data-movement optimization, i.e., sending the analysis code as close as
possible to the data. In practice, this means we focus on identifying sharding
of datasets and execute code accordingly over the appropriate shards.

@@

\begin{figure}[htbp]
\centering
\vspace{-0.10in}
\includegraphics[width=\columnwidth]{images/exa-arch.png}
\vspace{-0.20in}
\caption{Extreme Scale Architecture}
\label{fig:exa-arch}
\vspace{-0.10in}
\end{figure}

A common configuration for this type of deployment is shown in
Figure~\ref{fig:exa-arch}. The designated IO nodes (IONs) are connected to the
compute nodes (CNs) through the same fast fabric (e.g., InfiniBand) while the
connection to the external storage cluster is through a secondary, slower
channel (e.g., 10Gb Ethernet). By providing additional storage on the IO nodes,
such as SSDs, these nodes are capable of better regulating the IO pressure on
the underlying storage array better than simple forwarding gateways. For this
project, using something like SSDs on the IO nodes is termed a {\em Burst
Buffer} and is discussed below.

@@

The idea of burst buffers were initially explored in the context of data
staging~\cite{abbasi:2007:datatap,Abbasi:2009:datatap,nisar:2008:staging,zheng:2010:predata}.
These initial designs all use extra compute nodes to represent the data storage
buffer given the lack of any dedicated hardware support for this functionality.
The desired outcome of these initial studies is to motivate how such
functionality might be incorporated and the potential benefits.  Later, these
concepts were proposed to be incorporated into the existing IO stack
architecture~\cite{nowoczynski:2008:zest,bent:2012:challenges,bent:2012:burst-buffer}.

@@

@@

@@

@@

This is the layer where the container/object model is translated into the
physical storage requirements dictated by the physical storage underneath (the
VOSD layer). The two key design elements of this layer are the handling of
epochs and the mapping of containers and objects to the underlying storage.

@@

The current implementation has the DAOS layer map the container/object data
model onto a directory/file data model used for most existing file systems.
Should a fully object-based file system be deployed at the VOSD layer, this
mapping would be unnecessary. The current projections suggest that a standard
POSIX-like file system will likely be used at the lowest level on each storage
device requiring the mapping at some level. To perform this mapping, DAOS
considers the following.

@@

@@

At a more detailed level, the design for VOSD is an increment beyond the
current Lustre Object Storage Device design to incorporate the idea of shards
and the versioning aspects of transactions/epochs. For every DAOS shard, the
VOSD has information for storing and accessing the currently committed version,
the Highest Committed Epoch, as well as a staging dataset representing the
next version of the object being stored. Both of these are combined in a shard
root.

For data integrity, an intent log is maintained as part of the underlying file
system enabling fault recovery.

@@

In the subsections that follow, we examine transactions and epochs, metadata
management, function and analysis shipping, and the arbitrary connected graphs
support.

@@

As mentioned above, the transaction mechanism manifests in two forms. From the
user level down through the IOD layer, they are called transactions and are
used to judge whether or not a set of distributed, asynchronous modifications
across a set of related objects is complete or not.  It is also used to control
access by treating the transaction ID of committed transaction as a version
identifier.  At the DAOS layer and below, they are called epochs and represent
persisted (durable) transactions from the IOD layer. Each of these offers
different functionality, but are connected as is explained below.

@@

\subsubsection{Epochs}
The Epoch mechanism differs from transactions. Instead of focusing on when a
particular output is complete, an epoch represents incremental persisted
container copies.  To simplify the mapping between an IOD transaction and the
DAOS epochs, when an IOD transaction is persisted to DAOS, the IOD transaction
ID is the used as the epoch ID. The key difference is that at the DAOS layer,
some transaction (epoch) IDs will not be represented with data since not all
IOD transactions are necessarily persisted. Maintaining this ID continuity is
critical for multiplatform use. Since the shared point is the DAOS layer, any
user adding a new version to a file must be able to determine the most recent
transaction ID no matter from where the container was updated last.

\subsection{Metadata Management}
Metadata management has been a perennial challenge for parallel storage
systems.  Eliminating metadata management as a special case and instead
treating it just as data is a central design goal of the Fast Forward project.
This is a hybrid approach to metadata management that is halfway between
providing no inherent metadata support and having a fully integrated, but
separate metadata management system.

@@
